---
title: "Reproducible Research: Peer Assessment 1"
author: "ckwong338"
date: "Monday, October 20, 2014"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data

```{r}
activity <- read.csv("activity.csv")
activity$date <- as.Date(activity$date, "%Y-%m-%d")
```

## What is mean total number of steps taken per day?

```{r}
aggregated <- aggregate(steps ~ date, activity, sum)
```
```{r, echo=FALSE}
library(ggplot2)
ggplot(data=aggregated) + geom_histogram(aes(x = steps), binwidth = 2500, colour="green", fill="black") + labs(title="Total number of steps taken per day", x="Steps/Day", y="Count")
```

Mean of total number of steps taken per day:
```{r}
mean(aggregated$steps)
```
Median of total number of steps taken per day:
```{r}
median(aggregated$steps)
```
## What is the average daily activity pattern?

Apply mean across the 5-minute intervals:
```{r}
ts <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)
```
```{r, echo=FALSE}
plot(row.names(ts), ts, type="l", col="blue", xlab="5-minute interval",ylab="Average across all days", main="Time series plot of 5-minute intervals")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
names(ts[ts==max(ts)])
```

## Inputing missing values
Total number of missing values in the dataset:
```{r}
sum(is.na(activity$steps))
```
Fill in all missing values with mean for the 5-minute interval, and create new dataset:
```{r}
averagePerInterval <- aggregate(steps ~ interval, activity, mean)
newActivity <- activity
for (i in 1:nrow(newActivity)) {
    if (is.na(newActivity[i,]$steps)) {
        newActivity[i,]$steps<-averagePerInterval[averagePerInterval$interval==newActivity[i,]$interval,]$steps
    }
}
```
```{r}
aggregated2 <- aggregate(steps ~ date, newActivity, sum)
```
```{r, echo=FALSE}
library(ggplot2)
ggplot(data=aggregated2) + geom_histogram(aes(x = steps), binwidth = 2500, colour="green", fill="black") + labs(title="Total number of steps taken per day", x="Steps/Day", y="Count")
```

Mean of total number of steps taken per day:
```{r}
mean(aggregated2$steps)
```
Median of total number of steps taken per day:
```{r}
median(aggregated2$steps)
```
There is not much differences at all in the mean and median of the total number of steps taken per day.

## Are there differences in activity patterns between weekdays and weekends?

Create new factor variable for dataset, indicating weekday or weekend:
```{r}
dayFactor <- vector()
for (i in 1:nrow(newActivity)) {
    if (weekdays(newActivity[i,]$date) == "Saturday") {
        dayFactor[i] <- "Weekend"
    } else if (weekdays(newActivity[i,]$date) == "Sunday") {
        dayFactor[i] <- "Weekend"
    } else {
        dayFactor[i] <- "Weekday"
    }
}
newActivity$dayFactor <- factor(dayFactor)
aggregated3 <- aggregate(steps ~ interval + dayFactor, data = newActivity, mean)
```
```{r, echo=FALSE}
library(lattice)
xyplot(steps ~ interval | dayFactor, aggregated3, type = "l", layout = c(1, 2), 
    xlab = "Interval", ylab = "Number of steps")
```

Yes, there are differences in activity patterns between weekdays and weekends.
